---
title: "STAT 437 Project"
output: html_notebook
---

```{r}
library(lme4)
library(dplyr)

set.seed(123)
```

```{r}
mydata <- read.csv("/Users/aliyahaas/Downloads/STAT437_data.csv")
head(mydata)

#Change Readmission and Outcome to 0 and 1
mydata <- mydata %>% mutate(Readmission = ifelse(Readmission == "No",0,1))

#Outcome, 0 = Stable, 1 = Recovered
mydata <- mydata %>% mutate(Outcome = ifelse(Outcome == "Stable",0,1))

mydata$Condition <- as.factor(mydata$Condition)
mydata$Patient_ID <- as.factor(mydata$Patient_ID)
```


## Stepwise Selection
```{r}

# Fit the initial full model
initial_model <- lm(Satisfaction ~ ., data = mydata)

# Perform stepwise selection
stepwise_model <- step(initial_model, direction = "both")

# Print the summary of the selected model
summary(stepwise_model)
table(mydata$Condition, mydata$Procedure)

```
Satisfaction ~ Age + Condition
Significant p-values
Residual standard error: 0.1753 on 968 degrees of freedom
Multiple R-squared:  0.9612,	Adjusted R-squared:  0.9606 
F-statistic:  1598 on 15 and 968 DF,  p-value: < 2.2e-16


family=poisson(link="log")


## GLMM

```{r}
library(lme4)

# Condition = random effect
glmm_satisfaction <- lmer(Satisfaction ~ Age + (1 | Condition),
                           data = mydata,
                           REML = FALSE)

# Condition/Procedure = random effect (collinearity)
glmm_satisfaction <- lmer(Satisfaction ~ Age + (1 | Condition/Procedure),
                           data = mydata,
                           REML = FALSE)

summary(glmm_satisfaction)


```

















## Drawing Board (just exploring the data)

```{r}

model <- lm(Satisfaction ~ Age + Condition, data = mydata)
summary(model)


```

In the paper, they used GLMM to study the general effects of elevation, flower time, and floral morph

I can study the general effects of Cost, Length_of_Stay, and Condition on Satisfaction

What would be a Fixed Effect? Random Effect?


I could also look at the general effects of Procedure, Cost, and Length_of_stay on Outcome (1 or 0)


normal distribution and log-linked function



```{r}
hist(mydata$Cost)
library(MASS)
boxcox(Cost ~ Age + Condition, data = mydata)

mydata$Cost_bc <- mydata$Cost^0.6
hist(mydata$Cost_bc)

model_bc <- lm(Cost_bc ~ Age + Condition, data = mydata)

plot(model_bc)
```



