---
title: "Stat project"
author: "Elena, Olive, Aliya"
date: "2026-03-31"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import and loading libraries
```{r}
set.seed(99999)
library(dplyr)
library(car)
#change path according to where you have the data set stored
Hospital_Data <- read.csv("/Users/eepolonsky/Desktop/Stat\ 437/Project/heart_failure_clinical_records_dataset.csv")
```

# Convert binary variables into labeled factors
```{r}
Hospital_Data$DEATH_EVENT <- factor(Hospital_Data$DEATH_EVENT, 
                                    levels = c(0, 1), 
                                    labels = c("Survived", "Died"))

Hospital_Data$diabetes <- factor(Hospital_Data$diabetes, 
                                levels = c(0, 1), 
                                labels = c("No_Diabetes", "Diabetes"))

Hospital_Data$anaemia <- factor(Hospital_Data$anaemia, 
                                levels = c(0, 1), 
                                labels = c("No_Anaemia", "Anaemia"))

Hospital_Data$high_blood_pressure <- factor(Hospital_Data$high_blood_pressure, 
                                levels = c(0, 1), 
                                labels = c("No_Hypertension", "Hypertension"))

Hospital_Data$sex <- factor(Hospital_Data$sex, 
                                levels = c(0, 1), 
                                labels = c("Female", "Male"))

Hospital_Data$smoking <- factor(Hospital_Data$smoking, 
                                levels = c(0, 1), 
                                labels = c("Non_Smoker", "Smoker"))

```

# Define continuous outcome variables and grouping variables
```{r}
continuous_vars <- c(
  "age",
  "creatinine_phosphokinase",
  "ejection_fraction",
  "platelets",
  "serum_creatinine",
  "serum_sodium",
  "time"
)

grouping_vars <- c(
  "DEATH_EVENT",
  "anaemia",
  "diabetes",
  "high_blood_pressure",
  "sex",
  "smoking"
)
```

# Function to run Shapiro-Wilk test and return results in a single-row table
```{r}
run_shapiro <- function(data, var) {
  
  x <- data[[var]]
  x <- x[!is.na(x)]  # remove NA values
  
  test <- shapiro.test(x)
  
  data.frame(
    Variable = var,
    W_statistic = test$statistic,
    p_value = test$p.value,
    Normal = ifelse(test$p.value > 0.05, "Yes", "No")
  )
}
```

#  Loop to run all the Shapiro-Wilk tests
```{r}
shapiro_results_list <- list()

for (v in continuous_vars) {
  shapiro_results_list[[length(shapiro_results_list) + 1]] <- run_shapiro(Hospital_Data, v)
}

shapiro_results <- do.call(rbind, shapiro_results_list)
shapiro_results
```

Shapiro-Wilk tests were used to assess normality of continuous variables. All continuous variables significantly deviate from normality. Nonparametric Wilcoxon rank-sum tests were used instead of t-tests.

# Function to run Wilcoxon rank-sum tests and return results in a single-row table
```{r}
run_wilcox_table <- function(data, variable, grouping_variable) {
  
  form <- as.formula(paste(variable, "~", grouping_variable))
  test <- wilcox.test(form, data = data)
  
  p <- test$p.value
  
  data.frame(
    Test = paste(variable, "by", grouping_variable),
    p_value = p,
    Significant = ifelse(p < 0.05, "Yes", "No")
  )
}
```

# Loop to run all the Wicox tests for all combinations of continuous and grouping variables
```{r}
results_list <- list()

for (g in grouping_vars) {
  for (y in continuous_vars) {
    results_list[[length(results_list) + 1]] <- 
      run_wilcox_table(Hospital_Data, y, g)
  }
}

results <- do.call(rbind, results_list)
results
```

# Filter statistically significant results and sort by p-value
```{r}
significant_results <- results[results$p_value < 0.05, ]

significant_results <- significant_results[
  order(significant_results$p_value),
]
significant_results
```

# Compute group medians for all statistically significant tests
```{r}
get_medians_from_test <- function(data, test_name) {
  
  parts <- strsplit(test_name, " by ")[[1]]
  
  variable <- parts[1]
  grouping_variable <- parts[2]
  
  medians <- tapply(data[[variable]], data[[grouping_variable]], median, na.rm = TRUE)
  
  median_text <- paste(
    paste0(names(medians), " = ", round(medians, 2)),
    collapse = ", "
  )
  
  data.frame(
    Test = test_name,
    Medians = median_text
  )
}

median_list <- lapply(significant_results$Test, function(x) {
  get_medians_from_test(Hospital_Data, x)
})

median_table <- do.call(rbind, median_list)
median_table
```

There are no population-level groupings in this dataset, only individual patients. Instead of computing the coefficient of variation and running t-tests, we did Levene's tests to assess differences in variance between groups. 

# Function to run Levene's tests and return results in a single-row table
```{r}
run_levene_table <- function(data, variable, grouping_variable) {
  
  form <- as.formula(paste(variable, "~", grouping_variable))
  
  test <- car::leveneTest(form, data = data, center = median)
  
  p <- test[1, "Pr(>F)"]
  
  data.frame(
    Test = paste(variable, "by", grouping_variable),
    p_value = p,
    Significant = ifelse(p < 0.05, "Yes", "No")
  )
}
```

# Loop to run all the Levene's tests for all combinations of continuous and grouping variables
```{r}
levene_results_list <- list()

for (g in grouping_vars) {
  for (y in continuous_vars) {
    levene_results_list[[length(levene_results_list) + 1]] <- 
      run_levene_table(Hospital_Data, y, g)
  }
}

levene_results <- do.call(rbind, levene_results_list)
levene_results
```

# Filter statistically significant results and sort by p-value
```{r}
significant_levene <- levene_results[levene_results$p_value < 0.05, ]

significant_levene <- significant_levene[
  order(significant_levene$p_value),
]

significant_levene
```


