---
title: "quant project"
output: word_document
date: "2026-04-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. linear model with normal distribution. (follow-up time) ~ (sex) + (anemia status)
```{r}
heart<-read.csv("C:/Users/Missm/Downloads/quant bioi/heartfailureCSV.csv", header = TRUE)
dim(heart)
head(heart)

#convert all variables to numeric
indx <- sapply(heart, is.integer)
heart[indx] <- lapply(heart[indx], function(x) as.numeric(as.character(x)))

#convert age.group variable to factor
heart$age.group <- as.factor(heart$age.group)

str(heart)

#create linear model
model1 <- lm(time ~ male + anaemia, data=heart)
summary(model1)
```
Male patients were less likely (B1 = -4.752 ± 9.378) to have a long follow-up period (p=0.6128, p>0.05, NOT significant).
Patients with anemia were less likely (B2 = -22.558 ± 9.039) to have a long follow-up period (p=0.0131, p<0.05, significant).

Sex is not a significant predictor of follow-up time period, but anemia status is.

y = 143.076 - 4.752(x1) - 22.558(x2)


2. logit-link model with binomial distribution. (death event) ~ (age group)
```{r}
model2 <- glm(death ~ (age.group - 1), family=binomial, data=heart) #intercept-free model to compare multiple categories
summary(model2)
exp(model2$coefficients)
```
Patients in age group 1 were less likely (B1 = -1.2321 ± 0.2480) to die during treatment (p=0.000000676, p<<<<0.05, very significant).
Patients in age group 2 were less likely (B2 = -0.9808 ± 0.2041) to die during treatment (p=0.00000155, p<<<<0.05, very significant).
Patients in age group 3 were less likely (B3 = -0.2963 ± 0.2452) to die during treatment (p=0.2269, p>>0.05, NOT significant).
Patients in age group 4 were more likely (B4 = 1.1787 ± 0.5718) to die during treatment (p=0.0393, p<0.05, significant).

fitted model: ln(p/1-p) = -1.2321(x1) - 0.9808(x2) - 0.2963(x3) + 1.1787(x4)

